Assessing the Accessibility of Protein Tunnels using Motion Planning

### 2017-12-07 paper.tex
last two paragraphs of experiments moved to discussion

In conclusion, the traversability of the DCP ligand to active site was approved by the MD simulation.
The observed trajectory of the DCP ligand in the MD simulation followed the main tunnel, as expected.
Based on the traversability rates with $\smin=0.8$, we expect that at least ligands m037t and m038t can also reach the active site, as
they have the same (or even larger) traversability rate than the DCP.
However, this hypothesis has not been approved yet by testing it on MD simulations.

@@

\vspace{-3mm}
The most time-consuming part of the planners is the collision detection. 
Although both methods can utilize the same number of collision detection queries in the expansion step, \RB\ is slower than \RA.
The reason is that \RA\ expands the tree more closer to the random samples $\qrand$ than \RB.
This results in a faster exploration of the configuration space and therefore, \RA\ needs less number of iterations to reach the active site.
The better performance of \RA\ is caused by the proposed atom-based metric, because this distance depends on the actual shape of the ligand and 
it allows the ligands to retract more towards $\qrand$.
On the contrary, \RB\ utilizes the single 6D Euclidean weighted metric in the expansion, where the weights are the same for all conformations.
Depending on these weights, the metric favors samples that differ from $\qrand$ more in the rotation or translation part.
However, the weights between the rotation/translation parts should depend on the shape of the ligand, which is different in each conformation.
By using a single set of weights in the 6D Euclidean metric, we implicitly favor either the packed or the longer conformations.

@@

Judging the accessibility of the active site by a ligand based only on the traversability rates may be problematic if the ligand is scaled too much.
Instead of using a single version of the scaled-down ligand, we only limit the minimal scale used by the planner (the parameter $\smin$).
The expansion step always prefers to expand the tree by a collision-free configuration with the maximal scale.
%that leads to a new collision-free configuration.
The information about the scales used at the trajectory points (which can be extracted from the configuration tree of the proposed planner) brings the biochemists new knowledge about the tunnels, which was impossible to get using the previous solutions.
Our proposed visualization allows the biochemists to easily observe those parts of the tunnel where the ligand had to be scaled-down significantly (red parts of trajectories in Fig.~\ref{fig::dcp}), and parts where the ligand could be enlarged again (green parts in Fig.~\ref{fig::dcp}).
According to the results of the motion planing, the DCP needs to shrink down to ca. 75~\% of its size (Fig.~\ref{fig::dcp}, point B).
However, as both the protein and ligand are mutually influence each other, it is possible that even a very narrow part of the tunnel is traversable.
This was observed in the MD simulation of the DCP ligand in the 1CQW protein, which revealed that DCP can pass the main tunnel.

>>>>

In conclusion, the traversability of the DCP ligand to active site was approved by the MD simulation~\cite{marques2017catalytic}.
The observed trajectory of the DCP ligand in the MD simulation followed the main tunnel, as expected.
Based on the traversability rates with $\smin=0.8$, we expect that at least ligands m037t and m038t can also reach the active site, as
they have the same (or even larger) traversability rate than the DCP.
However, this hypothesis has not been approved yet by testing it on MD simulations.

@@

@@

Judging the accessibility of the active site by a ligand based only on the traversability rates may be problematic if the ligand is scaled too much.
Instead of using a single version of the scaled-down ligand, we only limit the minimal scale used by the planner (the parameter $\smin$).
The expansion step always prefers to expand the tree by a collision-free configuration with the maximal scale.
%that leads to a new collision-free configuration.
The information about the scales used at the trajectory points (which can be extracted from the configuration tree of the proposed planner) brings the biochemists new knowledge about the tunnels, which was impossible to get using the previous solutions.
Our proposed visualization allows the biochemists to easily observe those parts of the tunnel where the ligand had to be scaled-down significantly (red parts of trajectories in Fig.~\ref{fig::dcp}), and parts where the ligand could be enlarged again (green parts in Fig.~\ref{fig::dcp}).
According to the results of the motion planing, the DCP needs to shrink down to ca. 75~\% of its size (Fig.~\ref{fig::dcp}, point B).
However, as both the protein and ligand are mutually influence each other, it is possible that even a very narrow part of the tunnel is traversable.
This was observed in the MD simulation of the DCP ligand in the 1CQW protein, which revealed that DCP can pass the main tunnel~\cite{marques2017catalytic}.

%originally last two paragraphs in the experiment section, for resubmition, they are moved here to discussion

The most time-consuming part of the planners is the collision detection. 
Although both methods can utilize the same number of collision detection queries in the expansion step, \RB\ is slower than \RA.
The reason is that \RA\ expands the tree more closer to the random samples $\qrand$ than \RB.
This results in a faster exploration of the configuration space and therefore, \RA\ needs less number of iterations to reach the active site.
The better performance of \RA\ is caused by the proposed atom-based metric, because this distance depends on the actual shape of the ligand and 
it allows the ligands to retract more towards $\qrand$.
On the contrary, \RB\ utilizes the single 6D Euclidean weighted metric in the expansion, where the weights are the same for all conformations.
Depending on these weights, the metric favors samples that differ from $\qrand$ more in the rotation or translation part.
However, the weights between the rotation/translation parts should depend on the shape of the ligand, which is different in each conformation.
By using a single set of weights in the 6D Euclidean metric, we implicitly favor either the packed or the longer conformations.
